Physcraper: A Python package for continually updated gene trees

### 2020-05-12 physcraper_ms.Rmd
doc version

@@

# Intro

@@

- The phylogenetic tree has to be in the Open Tree of Life store [@mctavish2015phylesystem].
  You can choose from a variety of published trees supporting any node of the Tree of Life. If the tree you are
  interested in is not in Open Tree of Life, you can easily upload it via the curator
  tool.
- The alignment should be a gene alignment that was used to generate the tree. The
  alignments are usually stored in a public repository such as TreeBase [@piel2009treebase; @vos2012nexml]
  or DRYAD (<http://datadryad.org/>). If the alignment is stored in TreeBase,
  `physcraper` can download it directly either from the TreeBASE website (<https://treebase.org/>)
  or through its GitHub repositiry, SuperTreeBASE (<https://github.com/TreeBASE/supertreebase>).
  If the alignment is on another repository, a copy of it has to be
  downloaded by the user, and it's location has to be provided as an argument.
- A taxon name matching step is performed to verify that all taxon names on the tips
  of the tree are in the character matrix and vice versa.
- A ".csv" file with the summary of taxon name matching is produced for the user.
- Unmatched taxon names are dropped from the tree and alignment. ***What is the minimum amount of tips necessary for a physcraper run??***
- A ".tre" and ".aln" files are generated and saved for a  `physcraper` run.

@@

- The first step is to identify the MRCA of all taxa in the tree that is also a
  named clade in the NCBI taxonomy. This "taxonomic MRCA" can be different from
  the phylogenetic MRCA when the latter is an unnamed clade. ***Does
  physcraper take into account the focal clade from Open Tree in any way???***
- The taxonomic MRCA is used to constrain the DNA sequence search within that taxonomic group on the GenBank database.
- The DNA sequence search can be done on a local database that is easily setup by the user, or remotely.
- The BLAST algorithm is used to identify all DNA sequences in the database within
  the taxonomic MRCA that are similar to each sequence in the alignment ***How does the blast is setup? Is it an all-blast-all??***
- Reverse complement sequences are identified and translated.
- A fasta file containing all sequences resulting from the search is generated for the user.

@@

- A gene tree is reconstructed for each alignment provided, using ***what software?***

>>>>

---
title: 'Physcraper: A Python package for continually updated gene trees'
date: 'Manuscript rendered on `r format(Sys.time(), "%d %B, %Y")`'
author:
  - Luna L. Sanchez Reyes, University of California, Merced
  - Martha Kandziora, University of California, Merced
  - Emily Jane McTavish, University of California, Merced
output:
  redoc::redoc
bibliography: paper.bib
csl: apa.csl
---

@@

# Introduction

@@

- The phylogenetic tree has to be in the Open Tree of Life store [@mctavish2015phylesystem].
  You can choose from a variety of published trees supporting any node of the Tree of Life. If the tree you are
  interested in is not in Open Tree of Life, you can easily upload it via the curator
  tool.
- The alignment should be a gene alignment that was used to generate the tree. The
  alignments are usually stored in a public repository such as TreeBase [@piel2009treebase; @vos2012nexml]
  or DRYAD (<http://datadryad.org/>). If the alignment is stored in TreeBase,
  `physcraper` can download it directly either from the TreeBASE website (<https://treebase.org/>)
  or through its GitHub repositiry, SuperTreeBASE (<https://github.com/TreeBASE/supertreebase>).
  If the alignment is on another repository, a copy of it has to be
  downloaded by the user, and it's location has to be provided as an argument.
- A taxon name matching step is performed to verify that all taxon names on the tips
  of the tree are in the character matrix and vice versa.
- A ".csv" file with the summary of taxon name matching is produced for the user.
- Unmatched taxon names are dropped from the tree and alignment.
  ***What is the minimum amount of tips necessary for a physcraper run??*** Just one!
- A ".tre" and ".aln" files are generated and saved for a  `physcraper` run.

@@

- The first step is to identify the MRCA of all taxa in the tree that is also a
  named clade in the NCBI taxonomy. This "taxonomic MRCA" can be different from
  the phylogenetic MRCA when the latter is an unnamed clade. ***Does
  physcraper take into account the focal clade from Open Tree in any way???*** Not really, it is a search MRCA more than a taxonomic, the default is to get the ingroup from Otol, which is not the focal group.
- The taxonomic MRCA is used to constrain the DNA sequence search within that taxonomic group on the GenBank database.
- The DNA sequence search can be done on a local database that is easily setup by the user, or remotely.
- The BLAST algorithm is used to identify all DNA sequences in the database within
  the taxonomic MRCA that are similar to each sequence in the alignment ***How does the blast is setup? Is it an all-blast-all??***
  The rmote BLAST uses bioPython and the local BLAST uses BLASTn
- Reverse complement sequences are identified and translated.
- Sequences found for each
- A fasta file containing all sequences resulting from the search is generated for the user.
- ***Add cycles of blast searches with new sequences***
- How many sequences per species? Was hardcoded to 5, but it will be an argument now, default to 5. ***How does pyphlawd, or phylota does it?***

@@

- A gene tree is reconstructed for each alignment provided, using RAxML with bootstrap replicates.
- The final result is a gene tree couples to the conlict info.
